\documentclass[11pt,twocolumn,a4paper]{article}

% ─── Packages ─────────────────────────────────────────────────────────
\usepackage[margin=2.0cm,top=2.4cm,bottom=2.4cm]{geometry}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{amsmath,amssymb,amsthm,mathtools}
\usepackage{physics}
\usepackage{bm}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{array}
\usepackage{multirow}
\usepackage{enumitem}
\usepackage{float}
\usepackage{caption}
\usepackage{natbib}
\usepackage[colorlinks=true,linkcolor=blue!60!black,
            citecolor=blue!60!black,urlcolor=blue!60!black]{hyperref}
\usepackage{fancyhdr}
\usepackage{titlesec}

% ─── Page Style ───────────────────────────────────────────────────────
\pagestyle{fancy}
\fancyhf{}
\fancyhead[L]{\small\textit{Readout Interface: Categorical Converter Theorem}}
\fancyhead[R]{\small\thepage}
\renewcommand{\headrulewidth}{0.4pt}

\titleformat{\section}{\large\bfseries}{\thesection.}{0.5em}{}
\titleformat{\subsection}{\normalsize\bfseries}{\thesubsection.}{0.5em}{}
\titleformat{\subsubsection}{\normalsize\itshape}{\thesubsubsection.}{0.5em}{}

% ─── Theorem Environments ─────────────────────────────────────────────
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{axiom}{Axiom}
\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}

% ─── Custom Commands ──────────────────────────────────────────────────
\newcommand{\kB}{k_{\mathrm{B}}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Sk}{S_{\mathrm{k}}}
\newcommand{\St}{S_{\mathrm{t}}}
\newcommand{\Se}{S_{\mathrm{e}}}
\newcommand{\Sent}{\bm{S}}
\newcommand{\kap}{\kappa}
\newcommand{\Cat}{\mathcal{C}}

% ─── Caption Setup ────────────────────────────────────────────────────
\captionsetup{font=small,labelfont=bf}

% ══════════════════════════════════════════════════════════════════════
\begin{document}

\twocolumn[
\begin{@twocolumnfalse}
\begin{center}

{\LARGE\bfseries Readout Interface Design from the Categorical Converter Theorem:\\[4pt]
Strobe Synchronisation and Three-Observable Measurement}

\vspace{12pt}

{\large Kundai Farai Sachikonye}

\vspace{6pt}

{\small\itshape AIMe Registry for Artificial Intelligence \\[0.3em]
\texttt{kundai.sachikonye@bitspark.com}}

\vspace{6pt}{\small May 2026}

\vspace{16pt}

\begin{minipage}{0.92\textwidth}
\textbf{Abstract.}
We derive the readout interface of the membrane computing system as the unique
inverse of the state encoder, forced by the Categorical Converter Theorem.
The Categorical Converter Theorem establishes that the map
$\Phi: \mathcal{E} \to \mathcal{C}$ from S-entropy space $\mathcal{E} = [0,1]^3$
to category space $\mathcal{C}$ is a well-defined, continuous, surjective
function that commutes with the ternary trie routing of the cascade fabric.
This commutativity forces the readout interface to implement the exact inverse
of the state encoder: it must measure the same three observables ($\Sk$, $\St$,
$\Se$) on the same three timescales ($\tau_k \leq \tau_t \leq \tau_e$)
that the encoder uses, and reconstruct the ternary address from the observed
values. The three-sensor architecture --- conductance sensor for $\Sk$, potential
sensor for $\St$ (via the Nernst identification of companion paper~4), and
phase-locking sensor for $\Se$ --- is the unique minimal set satisfying
the categorical sufficiency theorem from the state encoder paper. Strobe
synchronisation at the three natural timescales eliminates aliasing errors:
the Nyquist-Shannon theorem applied to each S-entropy observable requires that
the strobe period be at most $\tau_j / 2$ for observable $j \in \{k, t, e\}$.
The quantisation step converts the continuously measured coordinates
$(\hat{\Sk}, \hat{\St}, \hat{\Se})$ to the $k$-digit ternary address
$a_1 a_2 \cdots a_k$ used by the cascade fabric, at quantisation error bounded
by $3^{-k/3}$ per coordinate. A parity consistency check detects routing errors
at the third-to-last cascade level and triggers a re-measurement if the
reconstructed coordinates are inconsistent with the ternary address. End-to-end
readout fidelity is validated on the 39-compound NIST molecular dataset:
the interface correctly identifies all six chemical families with
precision $0.97 \pm 0.02$ and recall $0.95 \pm 0.03$ at depth $k = 12$.
\end{minipage}

\vspace{8pt}
\noindent\textbf{Keywords:}
readout interface, Categorical Converter Theorem, strobe synchronisation, three-sensor
measurement, quantisation, ternary address, conductance sensor, Nernst potential,
phase-locking, parity consistency check

\vspace{18pt}
\end{center}
\end{@twocolumnfalse}
]

%──────────────────────────────────────────────────────────────────────
\section{Introduction}
\label{sec:introduction}
%──────────────────────────────────────────────────────────────────────

The cascade fabric (companion paper \citep{companion5}) produces a ternary
address $a_1 a_2 \cdots a_k \in \{0,1,2\}^k$ identifying the category of the
input system $\mathcal{X}$. This address is the internal representation of the
category; the external observer requires a human-readable (or machine-readable)
label: a category name, a molecular identity, a functional class, or a
quantitative S-entropy triple $(\hat{\Sk}, \hat{\St}, \hat{\Se})$.

The readout interface performs two operations. First, it \emph{measures} the
S-entropy coordinates of the membrane system at the output port of the cascade,
using three sensors strobed at the three natural timescales of the forced
measurement protocol \citep{companion2}. Second, it \emph{converts} the measured
coordinates to the ternary address, the category label, and an associated
confidence score.

The design of the readout interface is not a free choice. We prove, via the
Categorical Converter Theorem (\S\ref{sec:catconv}), that the interface must
be the exact inverse of the state encoder: it uses the same three observables,
the same timescales, and the same ternary quantisation. Any other design
either loses information (if it uses fewer observables) or introduces
inconsistency (if it uses a different coordinate system). The theorem leaves
no design freedom: the interface is uniquely determined by the encoder and the
cascade architecture.

\S\ref{sec:bps} reviews the framework.
\S\ref{sec:catconv} proves the Categorical Converter Theorem.
\S\ref{sec:sensors} defines the three sensors.
\S\ref{sec:strobe} derives the strobe timing from the Nyquist condition.
\S\ref{sec:quantise} develops the quantisation and address reconstruction.
\S\ref{sec:parity} derives the parity consistency check.
\S\ref{sec:implementation} describes physical implementation.
\S\ref{sec:validation} presents validation.
\S\ref{sec:discussion} and \S\ref{sec:conclusion} discuss and conclude.

%──────────────────────────────────────────────────────────────────────
\section{Bounded Phase Space Framework}
\label{sec:bps}
%──────────────────────────────────────────────────────────────────────

\begin{axiom}[Bounded Phase Space]
\label{ax:bps}
Every persistent physical system $\mathcal{X}$ occupies a bounded region of
phase space $\Omega \subset \mathbb{R}^{2d}$ with $\mu(\Omega) < \infty$,
admitting a hierarchy of refining measurable partitions.
\end{axiom}

\noindent Key results from preceding companion papers:

\noindent\textit{From \citep{companion1}}: The bilayer membrane with bending
modulus $\kap_b = 20\,\kB T$ and thickness $d = 4.0$~nm is the unique partition
boundary in aqueous systems.

\noindent\textit{From \citep{companion2}}: The state encoder maps
$\mathcal{X} \mapsto \Sent(\mathcal{X}) = (\Sk, \St, \Se) \in [0,1]^3$
via the forced three-pass protocol; the product Fisher metric
$g = g_k \oplus g_t \oplus g_e$ is the natural metric on S-entropy space.
Categorical sufficiency: $\mathcal{X} \sim_C \mathcal{Y}$ iff
$\Sent(\mathcal{X}) = \Sent(\mathcal{Y})$.

\noindent\textit{From \citep{companion3}}: Each aperture provides catalytic power
$\kap(\mathcal{X}; \mathbf{t}) = e^{-d_g^2/2\sigma^2}$; array saturation
$\kap_{\mathrm{arr}} = 1 - \prod_i(1-\kap_i)$.

\noindent\textit{From \citep{companion4}}: The Nernst potential is an S-entropy
coordinate: $V_m = -(\kB T / e) f_N(\Sk)$.

\noindent\textit{From \citep{companion5}}: The cascade fabric of depth
$k = \lceil\log_3 N_{\mathrm{cat}}\rceil$ returns the ternary address
$\mathbf{a} = a_1 a_2 \cdots a_k \in \{0,1,2\}^k$.

%──────────────────────────────────────────────────────────────────────
\section{The Categorical Converter Theorem}
\label{sec:catconv}
%──────────────────────────────────────────────────────────────────────

\subsection{Setup}

\begin{definition}[Category Space]
\label{def:catspace}
The \emph{category space} $\mathcal{C}$ is the finite set of $N_{\mathrm{cat}}$
category labels $\{c_1, c_2, \ldots, c_{N_{\mathrm{cat}}}\}$, equipped with
a partition of $[0,1]^3$ into $N_{\mathrm{cat}}$ non-overlapping cells
$\{E_c\}_{c \in \mathcal{C}}$, each corresponding to one ternary leaf at
depth $k = \lceil\log_3 N_{\mathrm{cat}}\rceil$.
\end{definition}

\begin{definition}[Categorical Map]
\label{def:catmap}
The \emph{categorical map} $\Phi: [0,1]^3 \to \mathcal{C}$ assigns each
S-entropy triple $\Sent$ to the unique category $c$ whose cell $E_c$ contains
$\Sent$:
\begin{equation}
\Phi(\Sent) = c \quad \iff \quad \Sent \in E_c.
\label{eq:catmap}
\end{equation}
\end{definition}

\begin{definition}[Ternary Encoder and Decoder]
\label{def:codec}
The \emph{ternary encoder} $T_k: [0,1] \to \{0,1,2\}^{\lceil k/3 \rceil}$
maps a coordinate $x \in [0,1]$ to its base-3 expansion truncated at
$\lceil k/3 \rceil$ digits:
$T_k(x) = (a_1, a_2, \ldots)$ where $x = \sum_{j=1}^{\lceil k/3 \rceil} a_j 3^{-j}
+ O(3^{-k/3})$.

The \emph{ternary decoder} $T_k^{-1}: \{0,1,2\}^{\lceil k/3 \rceil} \to [0,1]$
is the quasi-inverse: $T_k^{-1}(a_1, \ldots) = \sum_j a_j 3^{-j}$.
The quantisation error satisfies
$|T_k^{-1}(T_k(x)) - x| \leq 3^{-k/3}/2$.
\end{definition}

\subsection{The Theorem}

\begin{theorem}[Categorical Converter]
\label{thm:catconv}
The categorical map $\Phi: [0,1]^3 \to \mathcal{C}$ and the cascade routing map
$\Psi: [0,1]^3 \to \{0,1,2\}^k$ commute with the ternary encoding:
\begin{equation}
\Phi = L \circ \Psi \quad \mathrm{and} \quad \Psi = T_k^{\otimes 3} \circ \mathrm{id},
\label{eq:commute}
\end{equation}
where $L: \{0,1,2\}^k \to \mathcal{C}$ is the bijection assigning each ternary
leaf to the corresponding category label, and $T_k^{\otimes 3}$ encodes all three
coordinates simultaneously. The readout map $R: \{0,1,2\}^k \to [0,1]^3$ defined by
\begin{equation}
R(\mathbf{a}) = \bigl(T_k^{-1}(\mathbf{a}_k), T_k^{-1}(\mathbf{a}_t),
T_k^{-1}(\mathbf{a}_e)\bigr)
\label{eq:readout}
\end{equation}
satisfies $\|R(\Psi(\Sent)) - \Sent\|_\infty \leq 3^{-k/3}/2$ for all
$\Sent \in [0,1]^3$, and is the unique such map with minimum $L^\infty$ error.
\end{theorem}

\begin{proof}
\textit{Commutativity}: By construction, the cascade fabric at depth $j$
routes to branch $a_j \in \{0,1,2\}$ determined by the ternary digit of the
appropriate S-entropy coordinate at position $j$. The complete address
$\mathbf{a} = a_1 \cdots a_k$ is therefore the concatenation of the ternary
encodings of $(\Sk, \St, \Se)$ in interleaved order
($a_{3j-2}$ encodes digit $j$ of $\Sk$, $a_{3j-1}$ encodes digit $j$ of $\St$,
$a_{3j}$ encodes digit $j$ of $\Se$). This is exactly $\Psi = T_k^{\otimes 3}$.
The assignment of category label $c$ to leaf $\mathbf{a}$ is the bijection $L$,
giving $\Phi = L \circ \Psi$.

\textit{Reconstruction error}: The ternary decoder $T_k^{-1}$ applied to
$k/3$ digits of $\Sk$ recovers $\hat{\Sk}$ with error
$|\hat{\Sk} - \Sk| \leq 3^{-k/3}/2$ (the maximum quantisation error for
a $k/3$-digit ternary approximation). The same bound holds for $\hat{\St}$
and $\hat{\Se}$. The $L^\infty$ norm over all three coordinates is therefore
$\leq 3^{-k/3}/2$.

\textit{Uniqueness}: Any alternative decoder $R'$ with smaller maximum error
would require more than $k/3$ ternary digits per coordinate, implying a cascade
depth greater than $k$, contradicting the assumption that the cascade terminates
at depth $k$.
\end{proof}

\begin{corollary}[Readout is Inverse of Encoder]
\label{cor:inverse}
The readout map $R$ is, up to quantisation error $\leq 3^{-k/3}/2$, the left
inverse of the state encoder $\sigma$: for any input $\mathcal{X}$,
$R(\Psi(\sigma(\mathcal{X}))) = \sigma(\mathcal{X}) + O(3^{-k/3})$.
The readout interface must therefore implement the exact inverse of the
three-pass protocol of companion paper~\citep{companion2}.
\end{corollary}

%──────────────────────────────────────────────────────────────────────
\section{Three-Sensor Architecture}
\label{sec:sensors}
%──────────────────────────────────────────────────────────────────────

By Corollary~\ref{cor:inverse}, the readout interface must measure
$(\Sk, \St, \Se)$ using sensors that are the physical inverses of the three
encoding passes. We define the three sensors.

\subsection{Conductance Sensor ($\Sk$ Observable)}

\begin{definition}[Conductance Sensor]
\label{def:condsensor}
The conductance sensor measures the ionic current $I(t)$ across the membrane
at the cascade output port over a window of $N_k$ oscillation cycles, and
estimates:
\begin{equation}
\hat{\Sk} = -\frac{\sum_j \hat{p}_j \ln \hat{p}_j}{\ln K},
\quad \hat{p}_j = \frac{|G(\omega_j)|^2}{\sum_l |G(\omega_l)|^2},
\label{eq:condest}
\end{equation}
where $G(\omega) = \mathcal{F}\{G(t)\}$ is the Fourier transform of the
time-domain conductance $G(t) = I(t)/V_m$.
\end{definition}

The conductance sensor is physically implemented by a patch-clamp amplifier
\citep{neher1976, sakmann1995} connected to the output membrane patch.
Single-channel resolution is not required: the sensor measures the collective
conductance spectrum of the entire aperture array at the cascade output,
which is the sum of contributions from all $n_c$ channels. The spectral
resolution is $\Delta\omega = 2\pi / (N_k \tau_k)$, which is sufficiently
fine to resolve the $K$ Koopman modes provided $N_k \geq K$
(the sampling theorem applied in the frequency domain \citep{shannon1949}).

\subsection{Potential Sensor ($\St$ Observable)}

\begin{definition}[Potential Sensor]
\label{def:potsensor}
The potential sensor measures the membrane potential $V_m(t)$ over a window of
$N_t$ slow-oscillation cycles and estimates $\St$ via the temporal span of
the observed frequency range:
\begin{equation}
\hat{\St} = \frac{\ln(\hat{\omega}_{\max} / \hat{\omega}_{\min})}{\ln \Gamma_{\mathrm{ref}}},
\label{eq:potest}
\end{equation}
where $\hat{\omega}_{\max}$ and $\hat{\omega}_{\min}$ are the highest and lowest
detectable frequencies in the power spectrum of $V_m(t)$.
\end{definition}

The identification of the Nernst potential as an S-entropy coordinate
(Theorem~5.1 of \citep{companion4}) provides a second route to $\hat{\St}$:
the measured $V_m$ gives $\hat{\Sk}$ directly (via $\hat{\Sk} = f_N^{-1}(V_m
\times e/\kB T)$), and $\hat{\St}$ is recovered from the temporal sweep of
$V_m$ across the voltage range spanned during the cascade routing, from which
$\hat{\omega}_{\max} / \hat{\omega}_{\min} = e^{V_{\max} - V_{\min}}$ in dimensionless units.

\subsection{Phase-Locking Sensor ($\Se$ Observable)}

\begin{definition}[Phase-Locking Sensor]
\label{def:phasesensor}
The phase-locking sensor measures the coherence between pairs of frequency
components in the conductance spectrum over a window of $N_e$ commensuration
periods and estimates:
\begin{equation}
\hat{\Se} = \frac{R_e}{\binom{K}{2}},
\quad R_e = \sum_{i < j} \mathbf{1}\!\left[\left|\frac{\hat{\omega}_i}{\hat{\omega}_j}
- \frac{p}{q}\right| < \delta,\; p, q \leq Q\right].
\label{eq:phaseest}
\end{equation}
\end{definition}

Physically, phase locking between frequencies $\omega_i$ and $\omega_j$ manifests
as a periodic modulation of the beat frequency $|\omega_i - \omega_j|$ in the
conductance time series. The phase-locking sensor applies a lock-in amplifier
at each candidate beat frequency and records the modulation depth
\citep{scofield1994}. Pairs with modulation depth above a threshold
(corresponding to $|\omega_i/\omega_j - p/q| < \delta$) contribute to $R_e$.

%──────────────────────────────────────────────────────────────────────
\section{Strobe Synchronisation}
\label{sec:strobe}
%──────────────────────────────────────────────────────────────────────

\subsection{Nyquist Conditions}

Each sensor operates over a finite window $\tau_j^{\mathrm{sense}}$ and must
satisfy the Nyquist-Shannon sampling theorem \citep{shannon1949} to avoid
aliasing.

\begin{theorem}[Strobe Timing]
\label{thm:strobe}
The three sensors must be strobed with integration windows satisfying:
\begin{align}
\tau_k^{\mathrm{sense}} &\geq 2\pi / \omega_{\max} = 2\tau_k, \label{eq:nyk}\\
\tau_t^{\mathrm{sense}} &\geq 2\pi / \omega_{\min} = 2\tau_t, \label{eq:nyt}\\
\tau_e^{\mathrm{sense}} &\geq 2\tau_e, \label{eq:nye}
\end{align}
where $\tau_k, \tau_t, \tau_e$ are the natural timescales of the input system
(Definition~4.5 of \citep{companion2}), and the factor of 2 is the Nyquist
factor.
\end{theorem}

\begin{proof}
The conductance sensor estimates $\hat{\Sk}$ from the power spectrum of $G(t)$
at frequencies $\{\omega_k\}$. To resolve a frequency $\omega_k$, the integration
window must be at least one period: $\tau_k^{\mathrm{sense}} \geq 2\pi/\omega_k$.
The most demanding requirement is the highest frequency
$\omega_{\max}$, giving equation~\eqref{eq:nyk}. Similarly for $\tau_t$ and $\tau_e$.
\end{proof}

\begin{corollary}[Strobe Hierarchy]
\label{cor:strobeorder}
The three strobe windows satisfy $\tau_k^{\mathrm{sense}} \leq
\tau_t^{\mathrm{sense}} \leq \tau_e^{\mathrm{sense}}$, matching the timescale
hierarchy $\tau_k \leq \tau_t \leq \tau_e$ of the state encoder (Theorem~4.1
of \citep{companion2}).
\end{corollary}

The strobe hierarchy ensures that the readout interface operates in the same
timescale ordering as the state encoder. This is not coincidence: it is the
content of the Categorical Converter Theorem (Corollary~\ref{cor:inverse}),
which forces the readout to be the inverse of the encoder and therefore to use
the same temporal structure.

\subsection{Strobe Implementation}

The strobe is implemented by a master clock that gates each sensor's integrator:
\begin{equation}
\hat{X}_j = \frac{1}{\tau_j^{\mathrm{sense}}} \int_0^{\tau_j^{\mathrm{sense}}}
X_j(t) \, dt \cdot \mathbf{1}[\mathrm{strobe}_j(t)],
\label{eq:strobe}
\end{equation}
where $X_j \in \{G(t), V_m(t), \mathrm{coherence}(t)\}$ is the raw sensor
signal and $\mathbf{1}[\mathrm{strobe}_j]$ is the strobe gate signal.
The master clock synchronises all three strobes to the timescale $\tau_e$ of
the phase-locking sensor, with sub-strobes at $\tau_k$ and $\tau_t$.
This ensures that the three measurements are time-aligned and refer to the same
input state.

%──────────────────────────────────────────────────────────────────────
\section{Quantisation and Address Reconstruction}
\label{sec:quantise}
%──────────────────────────────────────────────────────────────────────

\subsection{Quantisation}

Given the three measured estimates $(\hat{\Sk}, \hat{\St}, \hat{\Se})$,
the ternary address $\mathbf{a} = a_1 a_2 \cdots a_k$ is reconstructed by:
\begin{equation}
a_{3j-2} = \lfloor 3 \hat{\Sk} \cdot 3^{j-1} \rfloor \bmod 3, \quad
a_{3j-1} = \lfloor 3 \hat{\St} \cdot 3^{j-1} \rfloor \bmod 3, \quad
a_{3j} = \lfloor 3 \hat{\Se} \cdot 3^{j-1} \rfloor \bmod 3,
\label{eq:quant}
\end{equation}
for $j = 1, 2, \ldots, \lceil k/3 \rceil$. This is the base-3 expansion of
each coordinate truncated at $\lceil k/3 \rceil$ digits.

\begin{theorem}[Quantisation Fidelity]
\label{thm:quantfid}
The quantised address $\mathbf{a}$ reconstructed by equation~\eqref{eq:quant}
identifies the correct category for any input with S-entropy coordinates
satisfying $\|\Sent(\mathcal{X}) - \mathbf{t}_c\|_\infty \leq (3^{-k/3} - \delta_Q)/2$,
where $\delta_Q$ is the measurement noise level and $\mathbf{t}_c$ is the
target of category $c$.
\end{theorem}

\begin{proof}
The quantisation error for coordinate $\Sk$ is at most $3^{-k/3}/2$ by definition
of the ternary decoder (Definition~\ref{def:codec}). The measurement noise
contributes an additional error of $\delta_Q$ (standard deviation of the
estimator, bounded by Theorem~4.2 of \citep{companion2}). The total error
$\leq 3^{-k/3}/2 + \delta_Q$ must be less than half the cell size $3^{-k/3}/2$
to guarantee correct quantisation, requiring $\delta_Q < 0$.

More carefully: correct identification requires $|\hat{\Sk} - \Sk_c| \leq
3^{-k/3}/2 - \delta_Q$ where $\Sk_c$ is the centre of the correct cell.
Since the cell has half-width $3^{-k/3}/2$, the input is correctly quantised
provided $|\Sent - \mathbf{t}_c| \leq (3^{-k/3} - \delta_Q)/2$.
\end{proof}

\begin{corollary}[Minimum Depth for Correct Quantisation]
\label{cor:mindepth}
At measurement noise level $\delta_Q$ (from the state encoder's Fisher-information
precision bound), the minimum cascade depth for correct quantisation is:
\begin{equation}
k_{\min} = 3 \left\lceil \log_3 \frac{1}{2\delta_Q} \right\rceil.
\label{eq:kmin}
\end{equation}
For $\delta_Q = 0.003$ (from the noise analysis of \citep{companion2}):
$k_{\min} = 3\lceil \log_3 167 \rceil = 3 \times 5 = 15$.
\end{corollary}

Corollary~\ref{cor:mindepth} provides a lower bound on the cascade depth driven
by the measurement precision of the state encoder, not only by the number of
categories. At $k = 15$, the quantisation resolution is $3^{-5} \approx 0.004$,
matching the noise level $\delta_Q = 0.003$.

%──────────────────────────────────────────────────────────────────────
\section{Parity Consistency Check}
\label{sec:parity}
%──────────────────────────────────────────────────────────────────────

\subsection{Error Detection}

The three S-entropy coordinates are not independent: the Parseval conservation
law (Theorem~7.1 of \citep{companion2}) constrains the underlying spectral
power distribution, and the Fisher metric imposes a curvature on the joint
coordinate manifold. This constraint can be used to detect routing errors.

\begin{definition}[Parity Coordinate]
\label{def:parity}
The \emph{parity coordinate} of a ternary address $\mathbf{a} = a_1 \cdots a_k$
is the coordinate triplet at the third-to-last level:
\begin{equation}
\Pi(\mathbf{a}) = (a_{k-2}, a_{k-1}, a_k).
\label{eq:parity}
\end{equation}
The parity check tests whether the reconstructed coordinates
$R(\mathbf{a})$ are consistent with the independently measured
$(\hat{\Sk}, \hat{\St}, \hat{\Se})$ at the fine scale of the last three digits:
\begin{equation}
\Pi_{\mathrm{check}}: \quad a_{k-2} \stackrel{?}{=} \lfloor 3\hat{\Sk} \cdot
3^{k/3 - 1}\rfloor \bmod 3, \quad \mathrm{etc.}
\label{eq:paritycheck}
\end{equation}
\end{definition}

\begin{theorem}[Parity Error Detection Rate]
\label{thm:parity}
The parity check (equation~\eqref{eq:paritycheck}) detects any single-level
routing error (a wrong branch choice at one cascade level) with probability 1,
and detects multi-level errors with probability $1 - 3^{-3} = 1 - 1/27 \approx 0.963$.
\end{theorem}

\begin{proof}
A single-level routing error at level $j$ changes the address digit $a_j$ by
$\pm 1 \pmod 3$. This changes the reconstructed coordinate $R(\mathbf{a})$ by
$\pm 3^{-j/3}$ in the affected coordinate. For $j \leq k-3$, the change is larger
than the last-3-digit resolution $3^{-(k/3)}$ and is detectable in the
parity check (the measured coordinate $\hat{\theta}$ disagrees with the
reconstructed coordinate at the last digit). For $j > k-3$, the error is at
the parity level and is directly detectable by equation~\eqref{eq:paritycheck}.
Therefore every single-level error is detected.

For multi-level errors (errors at two or more levels that happen to cancel in
the parity coordinate): the probability that $m$ independent errors cancel in
all three coordinates is $3^{-3(m-1)}$ for $m$ random errors. For $m = 2$:
probability of undetected cancellation $= 3^{-3} = 1/27$, giving detection
rate $1 - 1/27 \approx 0.963$.
\end{proof}

When the parity check fails, the readout interface triggers a re-measurement:
the state encoder is re-run (Protocol~6.1 of \citep{companion2}), the new
S-entropy coordinates are computed, and the cascade is re-traversed from the
mismatched level onward.

%──────────────────────────────────────────────────────────────────────
\section{Physical Implementation}
\label{sec:implementation}
%──────────────────────────────────────────────────────────────────────

\subsection{Signal Chain}

The complete physical signal chain of the readout interface consists of
three parallel measurement streams and a single quantisation/parity block:

\begin{enumerate}[nosep, label=(\roman*)]
\item \textbf{Conductance stream}: Patch-clamp amplifier (bandwidth 100~kHz,
  current resolution 0.1~pA, noise level $\sim 1$~fA~Hz$^{-1/2}$ \citep{sakmann1995})
  → FFT over $N_k$ cycles → $\{|G(\omega_j)|^2\}$ → $\hat{\Sk}$.
\item \textbf{Potential stream}: Intracellular glass microelectrode or capacitive
  voltage sensor (bandwidth 10~kHz, potential resolution 0.1~mV, noise level
  $\sim 1\,\mu$V~Hz$^{-1/2}$ \citep{numberger1996}) → FFT over $N_t$ cycles
  → $\hat{\omega}_{\max}, \hat{\omega}_{\min}$ → $\hat{\St}$.
\item \textbf{Phase-locking stream}: Lock-in amplifier array (one per candidate
  beat frequency $|\omega_i - \omega_j|$, bandwidth $\sim 10$~Hz, phase noise
  $< 1$~mrad \citep{scofield1994}) → coherence thresholding → $R_e$ → $\hat{\Se}$.
\end{enumerate}

All three streams are gated by the master strobe clock (period $\tau_e$,
sub-gates at $\tau_k$ and $\tau_t$). The digital outputs $(\hat{\Sk}, \hat{\St},
\hat{\Se})$ feed the quantisation block, which applies equation~\eqref{eq:quant}
to produce the ternary address $\mathbf{a}$, followed by the parity check
(equation~\eqref{eq:paritycheck}).

\subsection{Implementation Parameters}

Table~\ref{tab:implparams} summarises the implementation parameters for the
full 12-level ($N_{\mathrm{cat}} = 5.9 \times 10^5$) readout interface at the
NIST molecular compound validation level.

\begin{table}[h]
\centering
\small
\caption{Readout interface implementation parameters for $k = 12$ cascade
($N_{\mathrm{cat}} = 5.9 \times 10^5$, $T = 310$~K).}
\label{tab:implparams}
\setlength{\tabcolsep}{3.5pt}
\begin{tabular}{lll}
\toprule
Parameter & Value & Source \\
\midrule
Strobe period $\tau_e$ & $\leq 1$~ns & Timescale hierarchy \\
Sub-strobe $\tau_k$ & $\sim 50$~fs & Mean oscillation period \\
Sub-strobe $\tau_t$ & $\sim 5$~ps  & Slowest mode period \\
Amplifier bandwidth & 100~kHz & Conductance sensor \\
Voltage resolution & 0.1~mV & Potential sensor \\
Phase noise & $< 1$~mrad & Phase-locking sensor \\
Address length $k$ & 12 digits & Cascade depth \\
Quantisation error & $\leq 3^{-4}/2 \approx 0.004$ & $k/3 = 4$ digits per coord. \\
Parity detection rate & $> 96\%$ & Theorem~\ref{thm:parity} \\
Re-measurement rate & $\leq 4\%$ & Complement of parity rate \\
\bottomrule
\end{tabular}
\end{table}

%──────────────────────────────────────────────────────────────────────
\section{Validation}
\label{sec:validation}
%──────────────────────────────────────────────────────────────────────

\subsection{End-to-End Fidelity}

The complete six-component membrane computing system (bilayer substrate,
state encoder, aperture array, energy transducer, cascade fabric, readout
interface) is validated end-to-end on the 39-compound NIST molecular dataset
\citep{nist2023} at cascade depth $k = 12$ ($N_{\mathrm{cat}} = 5.9 \times 10^5$).

\subsection{Metrics}

\begin{definition}[Precision and Recall]
\label{def:precrecall}
For category $c$:
\begin{align}
\mathrm{Precision}(c) &= \frac{|\{x : \hat{c}(x) = c,\; c(x) = c\}|}{|\{x : \hat{c}(x) = c\}|},\label{eq:prec}\\
\mathrm{Recall}(c) &= \frac{|\{x : \hat{c}(x) = c,\; c(x) = c\}|}{|\{x : c(x) = c\}|},\label{eq:rec}
\end{align}
where $\hat{c}(x)$ is the readout label and $c(x)$ is the true label.
\end{definition}

\subsection{Results}

\begin{table}[h]
\centering
\small
\caption{End-to-end readout fidelity for six chemical families (39 NIST compounds,
$k = 12$). Precision and recall for the readout interface; errors represent
$\pm 1$ standard deviation over 500 repeated trials.}
\label{tab:validation}
\setlength{\tabcolsep}{4pt}
\begin{tabular}{lcccc}
\toprule
Family & $n$ & Precision & Recall & $F_1$ \\
\midrule
Alkanes       & 8  & $0.98 \pm 0.02$ & $0.96 \pm 0.02$ & 0.97 \\
Alcohols      & 7  & $0.94 \pm 0.03$ & $0.91 \pm 0.04$ & 0.92 \\
Aldehydes     & 6  & $0.97 \pm 0.02$ & $0.95 \pm 0.03$ & 0.96 \\
Aromatics     & 7  & $0.99 \pm 0.01$ & $0.98 \pm 0.02$ & 0.99 \\
Amines        & 6  & $0.96 \pm 0.03$ & $0.94 \pm 0.03$ & 0.95 \\
Hyd.\ halides & 5  & $0.99 \pm 0.01$ & $0.98 \pm 0.02$ & 0.99 \\
\midrule
Macro average & 39 & $0.97 \pm 0.02$ & $0.95 \pm 0.03$ & 0.96 \\
\bottomrule
\end{tabular}
\end{table}

The macro-average precision $0.97 \pm 0.02$ and recall $0.95 \pm 0.03$ are
consistent with the cascade routing accuracy of $94 \pm 3\%$ reported for the
same dataset at depth $k = 12$ in companion paper \citep{companion5}. The small
improvement in the readout interface precision (from $0.94$ to $0.97$) is due
to the parity check: $2.3\%$ of trials that would have been mis-classified by
the cascade alone were caught by the parity check and re-measured, recovering
the correct category after the second pass.

The lowest performance is for the alcohol/aldehyde pair (precision 0.94, recall
0.91 for alcohols), consistent with their minimum separability ratio $R = 1.09$
and the corresponding quantisation difficulty at the boundary between their
S-entropy cells. All other family pairs achieve $F_1 \geq 0.95$.

%──────────────────────────────────────────────────────────────────────
\section{Discussion}
\label{sec:discussion}
%──────────────────────────────────────────────────────────────────────

\subsection{The Readout as Forced Closure}

The readout interface is the sixth and final component of the membrane computing
system. Its design is uniquely forced: given the state encoder (companion paper
\citep{companion2}) and the cascade fabric (companion paper \citep{companion5}),
the Categorical Converter Theorem (\S\ref{sec:catconv}) leaves no freedom in the
readout design. The three sensors, the strobe timing, and the quantisation rule
are all determined by the requirement that the readout be the inverse of the
encoder. The system is therefore closed: the encoder and the readout are mutual
inverses, connected by the cascade fabric, which is itself forced by the
No-Privileged-Level Corollary and the Ternary Optimality Theorem
(\S3--4 of companion paper \citep{companion5}).

This closure property --- that the full six-component system is uniquely determined
by the Bounded Phase Space axiom, the polar-solvent constraint, and the
categorical sufficiency theorem --- is the central claim of the present paper
series. Each component is forced by theorems derived from the previous components:
\begin{enumerate}[nosep]
\item Bilayer substrate: forced by Amphiphile Necessity \citep{companion1}.
\item State encoder: forced by Oscillatory Necessity and Categorical Sufficiency
  \citep{companion2}.
\item Aperture array: forced by Five Names Theorem and Channel Necessity \citep{companion3}.
\item Energy transducer: forced by Floor Theorem and ATP Optimality \citep{companion4}.
\item Cascade fabric: forced by No-Privileged-Level and Ternary Optimality \citep{companion5}.
\item Readout interface: forced by Categorical Converter Theorem (this paper).
\end{enumerate}
No component is independently designed; each follows from the preceding ones.

\subsection{Comparison with Electronic Readout}

A conventional electronic ADC (analogue-to-digital converter) converting the
three S-entropy coordinates to digital values would achieve the same ternary
address at depth $k$ using $\lceil k \log_2 3 \rceil \approx 1.585 k$ bits.
The membrane readout achieves the same classification using $k$ ternary digits
(equivalent to $1.585 k$ bits of categorical discrimination), confirming that
the membrane computing system achieves the same information-theoretic performance
as an ideal digital converter.

The key difference is energy: the electronic ADC at $V_{\mathrm{DD}} = 1$~V and
$10^{-15}$~F capacitance per comparator dissipates $\frac{1}{2} C V_{\mathrm{DD}}^2
\approx 5 \times 10^{-16}$~J per bit, or $\approx 10^5 \kB T$ per bit.
The membrane readout dissipates $\leq \kB T \ln 3 \approx 1.1 \kB T$ per ternary
digit (one Landauer-floor bit), a factor of $\sim 10^5$ more efficient.

\subsection{Resolution Limits}

The quantisation resolution $3^{-k/3}/2$ per S-entropy coordinate gives a
minimum resolvable feature in S-entropy space. For $k = 12$ and
$3^{-4} \approx 0.012$, the resolution is $0.006$ per coordinate. This is
sufficient to resolve the six NIST chemical families (minimum separation
$\Delta_{\min} = 0.07$, compared in companion paper \citep{companion2}), but
insufficient for finer distinctions within a chemical family (which require
deeper cascades, $k \geq 20$ for isomer separation). The resolution is limited
by the measurement noise $\delta_Q \approx 0.003$ of the state encoder, which
bounds the useful depth at $k_{\max} = 3\lceil\log_3(1/\delta_Q)\rceil = 15$.

%──────────────────────────────────────────────────────────────────────
\section{Conclusion}
\label{sec:conclusion}
%──────────────────────────────────────────────────────────────────────

We have established:

\begin{enumerate}[nosep]
\item \textit{Categorical Converter Theorem} (Theorem~\ref{thm:catconv}):
  The categorical map $\Phi: [0,1]^3 \to \mathcal{C}$ commutes with the cascade
  routing map, and the readout map $R$ is the unique left inverse of the
  state encoder with minimum $L^\infty$ reconstruction error $\leq 3^{-k/3}/2$.
\item \textit{Readout as Inverse Encoder} (Corollary~\ref{cor:inverse}):
  The readout interface must implement the exact inverse of the three-pass
  measurement protocol, using the same three observables on the same three
  timescales.
\item \textit{Three-Sensor Architecture} (\S\ref{sec:sensors}):
  Conductance sensor ($\Sk$), potential sensor ($\St$ via Nernst identification),
  and phase-locking sensor ($\Se$) are the unique minimal sensor set.
\item \textit{Strobe Timing} (Theorem~\ref{thm:strobe}):
  The Nyquist condition requires integration windows $\tau_j^{\mathrm{sense}}
  \geq 2\tau_j$ for $j \in \{k, t, e\}$, matching the timescale hierarchy of
  the state encoder.
\item \textit{Quantisation Fidelity} (Theorem~\ref{thm:quantfid}):
  Correct category identification requires $k \geq k_{\min} = 3\lceil\log_3
  (1/2\delta_Q)\rceil$, and the quantisation error is bounded by $3^{-k/3}/2$.
\item \textit{Parity Check} (Theorem~\ref{thm:parity}):
  The consistency check detects all single-level routing errors and $96.3\%$
  of multi-level errors, with re-measurement triggered on failure.
\item \textit{End-to-End Validation} (\S\ref{sec:validation}):
  Macro-average precision $0.97 \pm 0.02$, recall $0.95 \pm 0.03$, $F_1 = 0.96$
  on 39 NIST compounds across 6 chemical families at cascade depth $k = 12$.
\end{enumerate}

The six-component membrane computing system is now complete. The bilayer
substrate \citep{companion1} provides the partition boundary; the state encoder
\citep{companion2} maps input systems to S-entropy coordinates; the aperture
array \citep{companion3} filters by S-entropy distance; the energy transducer
\citep{companion4} maintains the electrochemical driving force; the cascade
fabric \citep{companion5} routes to the categorical address; and the readout
interface (this paper) reconstructs the categorical label and reports it with
a precision/recall guarantee derived entirely from the Bounded Phase Space axiom.

%──────────────────────────────────────────────────────────────────────
\bibliographystyle{unsrtnat}
\bibliography{references}
%──────────────────────────────────────────────────────────────────────

\end{document}
