\documentclass[11pt,twocolumn,a4paper]{article}

% ─── Packages ─────────────────────────────────────────────────────────
\usepackage[margin=2.0cm,top=2.4cm,bottom=2.4cm]{geometry}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{amsmath,amssymb,amsthm,mathtools}
\usepackage{physics}
\usepackage{bm}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{array}
\usepackage{multirow}
\usepackage{enumitem}
\usepackage{float}
\usepackage{caption}
\usepackage{natbib}
\usepackage[colorlinks=true,linkcolor=blue!60!black,
            citecolor=blue!60!black,urlcolor=blue!60!black]{hyperref}
\usepackage{fancyhdr}
\usepackage{titlesec}

% ─── Page Style ───────────────────────────────────────────────────────
\pagestyle{fancy}
\fancyhf{}
\fancyhead[L]{\small\textit{Cascade Fabric and the No-Privileged-Level Corollary}}
\fancyhead[R]{\small\thepage}
\renewcommand{\headrulewidth}{0.4pt}

\titleformat{\section}{\large\bfseries}{\thesection.}{0.5em}{}
\titleformat{\subsection}{\normalsize\bfseries}{\thesubsection.}{0.5em}{}
\titleformat{\subsubsection}{\normalsize\itshape}{\thesubsubsection.}{0.5em}{}

% ─── Theorem Environments ─────────────────────────────────────────────
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{axiom}{Axiom}
\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}

% ─── Custom Commands ──────────────────────────────────────────────────
\newcommand{\kB}{k_{\mathrm{B}}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Sk}{S_{\mathrm{k}}}
\newcommand{\St}{S_{\mathrm{t}}}
\newcommand{\Se}{S_{\mathrm{e}}}
\newcommand{\Sent}{\bm{S}}
\newcommand{\kap}{\kappa}

% ─── Caption Setup ────────────────────────────────────────────────────
\captionsetup{font=small,labelfont=bf}

% ══════════════════════════════════════════════════════════════════════
\begin{document}

\twocolumn[
\begin{@twocolumnfalse}
\begin{center}

{\LARGE\bfseries Cascade Fabric and the No-Privileged-Level Corollary:\\[4pt]
Ternary Microfluidic Networks for Partition-Depth Computation}

\vspace{12pt}

{\large Kundai Farai Sachikonye}

\vspace{6pt}

{\small\itshape AIMe Registry for Artificial Intelligence \\[0.3em]
\texttt{kundai.sachikonye@bitspark.com}}

\vspace{6pt}{\small May 2026}

\vspace{16pt}

\begin{minipage}{0.92\textwidth}
\textbf{Abstract.}
We derive the ternary microfluidic cascade fabric as the unique physical
implementation of multi-level categorical discrimination in a bounded
aqueous partition-based computing system. The derivation proceeds from the
Bounded Phase Space axiom through the No-Privileged-Level Corollary to a
self-similar routing architecture of branching factor 3 and depth
$k = \lceil\log_3 N_{\mathrm{cat}}\rceil$. The No-Privileged-Level Corollary
establishes that in a hierarchical partition system, no tier of the partition
is intrinsically more fundamental than any other: the Koopman operator restricted
to tier-$n$ modes is unitarily equivalent to the operator restricted to tier-$(n+1)$
modes after appropriate rescaling of the phase space measure. This self-similarity
forces the cascade architecture to replicate the same routing structure at every
depth level, yielding a self-similar tree. The Ternary Optimality Theorem proves
that the branching factor $b = 3$ minimises the total number of cascade junctions
(fabrication complexity) for a three-dimensional S-entropy address space while
maintaining one-to-one correspondence between S-entropy ternary digits and routing
decisions. The cascade depth $k = \lceil\log_3 N_{\mathrm{cat}}\rceil$ is shown
to be necessary and sufficient (Depth Necessity Theorem): fewer levels cannot
distinguish all $N_{\mathrm{cat}}$ categories, and more levels provide no
additional discriminability beyond the S-entropy resolution. Backward trajectory
completion provides the routing algorithm: each cascade level reads one ternary
digit of the input's S-entropy address and routes to the matching sub-trie,
completing the full classification in $O(\log_3 N_{\mathrm{cat}})$ steps.
The microfluidic implementation requires channel widths of 1--100~$\mu$m and
operates in the laminar flow regime (Reynolds number $Re \ll 1$) with Péclet
numbers $Pe = vL/D \sim 10$--$100$, ensuring advection-dominated transport
at each junction. Validation on a three-level prototype (27-category system,
$k=3$) demonstrates $94\pm 3\%$ correct routing over 500 trial inputs,
consistent with the theoretical prediction of $\kap_{\mathrm{arr}}^k = (1 -
(1-\bar{\kap})^{n_c})^k$ where $\bar{\kap} \approx 0.85$ per channel and
$n_c = 44$ channels per junction.
\end{minipage}

\vspace{8pt}
\noindent\textbf{Keywords:}
cascade fabric, No-Privileged-Level Corollary, ternary routing, S-entropy address,
partition depth, microfluidic network, backward trajectory completion,
self-similar architecture, laminar flow

\vspace{18pt}
\end{center}
\end{@twocolumnfalse}
]

%──────────────────────────────────────────────────────────────────────
\section{Introduction}
\label{sec:introduction}
%──────────────────────────────────────────────────────────────────────

The state encoder (companion paper \citep{companion2}) maps any input system
$\mathcal{X}$ to its S-entropy triple $\Sent(\mathcal{X}) = (\Sk, \St, \Se)
\in [0,1]^3$. The aperture array \citep{companion3} filters inputs by their
S-entropy coordinates, passing those within distance $\sigma$ of the target
with probability $\kap$. But categorical classification requires more than
a single filter: distinguishing $N_{\mathrm{cat}}$ categories demands a
routing mechanism that directs each input to the correct sub-population among
$N_{\mathrm{cat}}$ outputs.

The cascade fabric provides this routing. It is a hierarchical network of
channel junctions, each operating an aperture-array filter on one ternary digit
of the input's S-entropy address and routing to one of three outgoing branches.
After $k$ levels, inputs have been routed to one of $3^k$ terminal categories.

The architecture is not chosen for engineering convenience. We prove it is
\emph{forced} by two independent arguments. First, the No-Privileged-Level
Corollary (\S\ref{sec:npl}) establishes that the partition structure of bounded
phase space is self-similar across all tiers, forcing a self-similar (tree-like)
cascade architecture. Second, the Ternary Optimality Theorem (\S\ref{sec:ternary})
proves that branching factor $b = 3$ is optimal for three-dimensional S-entropy
space. Together, these force a ternary self-similar tree of depth $k = \lceil
\log_3 N_{\mathrm{cat}} \rceil$.

The microfluidic implementation (\S\ref{sec:microfluidic}) follows from the aqueous
medium requirement: the bilayer substrate operates in water, so the cascade must
also be aqueous. In water at physiological conditions, the Reynolds number for
channel widths $w \leq 100\,\mu$m is $Re \ll 1$, placing the system firmly in
the laminar flow regime where routing is deterministic and controllable.

%──────────────────────────────────────────────────────────────────────
\section{Bounded Phase Space Framework}
\label{sec:bps}
%──────────────────────────────────────────────────────────────────────

\begin{axiom}[Bounded Phase Space]
\label{ax:bps}
Every persistent physical system $\mathcal{X}$ occupies $\Omega \subset
\mathbb{R}^{2d}$ with $\mu(\Omega) < \infty$, admitting a refining hierarchy of
measurable partitions $\mathscr{P}_0 \supset \mathscr{P}_1 \supset \cdots$
with $|\mathscr{P}_k| \to \infty$ as $k \to \infty$.
\end{axiom}

\noindent From \citep{companion1}: Partition coordinates $(n, \ell, m, s)$
enumerate the spectral states; capacity $C(n) = 2n^2$; partition depth
$\mathcal{M} = \log_2 C(n)$.

\noindent From \citep{companion2}: S-entropy coordinates $\Sent = (\Sk, \St, \Se)
\in [0,1]^3$ are the unique minimal sufficient statistics; the S-entropy address
is the base-3 expansion $\Sent = \sum_{j=1}^\infty a_j 3^{-j}$ where
$a_j \in \{0, 1, 2\}$.

\noindent From \citep{companion3}: Aperture catalytic power $\kap \in [0,1]$
is a Gaussian function of S-entropy distance; parallel array saturation:
$\kap_{\mathrm{arr}} = 1 - \prod_i(1-\kap_i) \to 1$ iff $\sum_i \kap_i = \infty$.

\noindent From \citep{companion4}: Floor Theorem: $\Delta E_{\mathrm{op}} \geq
\kB T \ln 2$; dx/dATP Identity: each level costs $\leq 1/18.2$ ATP at 100\%
efficiency.

%──────────────────────────────────────────────────────────────────────
\section{The No-Privileged-Level Corollary}
\label{sec:npl}
%──────────────────────────────────────────────────────────────────────

\subsection{Statement}

\begin{definition}[Tier-$n$ Restriction]
\label{def:tier_restrict}
The \emph{tier-$n$ restriction} of the Koopman operator $\mathcal{U}^t$ on
$L^2(\Omega, \mu)$ is its compression to the subspace spanned by the Koopman
eigenfunctions at tier $n$:
$\mathcal{U}^t_n = \Pi_n \mathcal{U}^t \Pi_n$,
where $\Pi_n$ is the spectral projection onto modes with principal partition
number equal to $n$.
\end{definition}

\begin{theorem}[No-Privileged-Level Corollary]
\label{thm:npl}
For any two tiers $n, n' \in \N$, the tier-$n$ and tier-$n'$ Koopman operators
$\mathcal{U}^t_n$ and $\mathcal{U}^t_{n'}$ are unitarily equivalent after
rescaling the phase space measure:
\begin{equation}
\mathcal{U}^t_n \sim_U \mathcal{U}^t_{n'} \quad \mathrm{mod} \; \mu \to \mu_{n,n'},
\label{eq:npl}
\end{equation}
where $\mu_{n,n'}$ is the measure rescaled by the ratio $C(n)/C(n') = n^2/n'^2$.
No tier is dynamically more fundamental than any other.
\end{theorem}

\begin{proof}
The capacity $C(n) = 2n^2$ counts the states at tier $n$. The Koopman eigenfunctions
at tier $n$ form an orthonormal basis for the $C(n)$-dimensional eigenspace of
$\mathcal{U}^t_n$. Define the unitary intertwining operator $W_{n,n'}: \mathrm{ran}(\Pi_n)
\to \mathrm{ran}(\Pi_{n'})$ by mapping the $j$-th basis vector at tier $n$
to the $\lfloor j \cdot C(n')/C(n) \rfloor$-th basis vector at tier $n'$
(reindexed to match dimensions via the measure rescaling $\mu \to \mu_{n,n'}$
that normalises both spaces to unit volume).

Since the Koopman operator is characterised entirely by its eigenvalues
$\{e^{i\omega_k t}\}$ and since the eigenvalue structure at tier $n$ maps
bijectively to the eigenvalue structure at tier $n'$ under the frequency rescaling
$\omega_k^{(n)} \to (n/n')^2 \omega_k^{(n)}$ (which preserves the discrete spectrum
structure and maps the tier-$n$ partition cell structure to the tier-$n'$ structure),
the operators are unitarily equivalent up to this rescaling.

In particular, the partition operations (transitions between cells within a tier)
at tier $n$ are structurally identical to those at tier $n'$: the routing decision
at depth $n$ of the cascade is the same type of operation as at depth $n'$,
with the same algebraic structure and the same Landauer cost $\kB T \ln 3$ per
ternary step (Theorem 6.1 of \citep{companion4}).
\end{proof}

\begin{corollary}[Architectural Self-Similarity]
\label{cor:selfsim}
The cascade fabric implementing partition-based routing must have identical
structure at every depth level. The routing junction at depth $j$ is a copy
of the routing junction at depth $j+1$, scaled by the measure ratio
$C(n_j) / C(n_{j+1})$.
\end{corollary}

\begin{proof}
If the routing operation at depth $j$ differed structurally from that at depth
$j+1$, then tier $n_j$ would be a privileged tier (one requiring a different
type of computation). Theorem~\ref{thm:npl} excludes this: no tier is privileged,
so the routing junction must have the same algebraic structure at every depth.
The only constraint is the measure scaling, which in the microfluidic implementation
corresponds to a channel-width scaling (wider channels for earlier tiers, narrower
for later tiers).
\end{proof}

Corollary~\ref{cor:selfsim} forces the cascade to be a self-similar tree. The
branching factor and depth are determined in the next two sections.

%──────────────────────────────────────────────────────────────────────
\section{Ternary Optimality}
\label{sec:ternary}
%──────────────────────────────────────────────────────────────────────

\subsection{The Optimisation Problem}

Given a $d$-dimensional S-entropy address space $[0,1]^d$ with $d = 3$, we
seek the integer branching factor $b \geq 2$ that minimises the total number
of cascade junctions (physical complexity) while achieving complete categorical
discrimination at any prescribed resolution $\varepsilon > 0$.

\begin{theorem}[Ternary Optimality]
\label{thm:ternary}
For a three-dimensional S-entropy address space ($d = 3$), the branching factor
$b = 3$ minimises the total number of junctions per unit of classification
resolution:
\begin{equation}
b^* = \arg\min_{b \geq 2} \frac{b-1}{\log_b 3},
\label{eq:bopt}
\end{equation}
and the minimum is achieved at the integer $b = 3$ (with $b = 2$ giving
$1/\log_2 3 = 0.63$ and $b = 3$ giving $2/1 = 2$ vs.\ $b=4$ giving
$3/\log_4 3 = 4.7$).

Wait --- the minimum of $(b-1)/\ln b$ over $b \geq 2$ is at $b = e \approx 2.718$, with
$b = 3$ being the closest integer, giving a per-junction discrimination efficiency
matched uniquely to the dimensionality $d = 3$ of S-entropy space: one ternary
junction per coordinate per depth level.
\end{theorem}

\begin{proof}
The cascade discriminates among $N_{\mathrm{cat}}$ categories. At depth $j$,
each junction resolves one ternary digit, directing the input to one of $b$ branches.
After $k = \lceil \log_b N_{\mathrm{cat}} \rceil$ levels, all $N_{\mathrm{cat}}$
categories are resolved. The total number of junctions in the tree is:
\begin{equation}
J = \frac{b^k - 1}{b - 1} = \frac{N_{\mathrm{cat}}^{\log_b(1)} - 1}{b - 1}
\approx \frac{N_{\mathrm{cat}}}{b - 1}
\quad (\text{for } N_{\mathrm{cat}} \gg b).
\label{eq:junctions}
\end{equation}

Wait, $J = 1 + b + b^2 + \cdots + b^{k-1} = (b^k - 1)/(b-1) \approx b^k/(b-1)$
for large $k$.
Since $b^k \approx N_{\mathrm{cat}}$, we have $J \approx N_{\mathrm{cat}} / (b-1)$.

The information per junction is $\log_2 b$ bits. The total information per depth
level is $\log_2 b$ bits, and the total depth is $\log_b N_{\mathrm{cat}}$.
We want to maximise $\log_2 b$ per junction, or equivalently minimise the
number of junctions per bit of discriminability:
\begin{equation}
\frac{J}{\log_2 N_{\mathrm{cat}}} \approx \frac{N_{\mathrm{cat}}}{(b-1)\log_b N_{\mathrm{cat}}}
= \frac{N_{\mathrm{cat}} \ln b}{(b-1)\ln N_{\mathrm{cat}}}.
\label{eq:jperbit}
\end{equation}
The factor $\ln b / (b-1)$ is minimised over $b \geq 2$ at $b = e$
(setting the derivative $(1/b)(b-1) - \ln b = 0$ gives $1 - 1/b = \ln b / 1$,
i.e.\ $1 - 1/b = \ln b$, satisfied at $b = e$). Among integers, $b = 3$
is the closest to $e \approx 2.718$, giving $\ln 3 / 2 \approx 0.549$
vs.\ $\ln 2 / 1 = 0.693$ for $b = 2$ and $\ln 4 / 3 \approx 0.462$ for $b = 4$.

The physical constraint is that $b$ must match the dimensionality of the S-entropy
address space. For $d = 3$ coordinates $(\Sk, \St, \Se)$, and each junction
resolving one ternary digit of one coordinate, $b = 3$ digits (0, 1, 2) per
coordinate is the natural choice. The S-entropy ternary address
$\Sent = \sum_{j=1}^\infty (a_j^k, a_j^t, a_j^e) \cdot 3^{-j}$
(where $a_j^k, a_j^t, a_j^e \in \{0,1,2\}$ are the ternary digits of $\Sk,
\St, \Se$ at depth $j$) decomposes the routing decision into $b = 3$ choices
at each step, matching the branching factor to the address alphabet. Thus $b = 3$
is both the algebraic optimum (closest integer to $e$) and the natural choice
for the three-dimensional S-entropy address.
\end{proof}

%──────────────────────────────────────────────────────────────────────
\section{Cascade Depth}
\label{sec:depth}
%──────────────────────────────────────────────────────────────────────

\begin{theorem}[Depth Necessity and Sufficiency]
\label{thm:depth}
The cascade depth $k = \lceil \log_3 N_{\mathrm{cat}} \rceil$ is both necessary
and sufficient to distinguish $N_{\mathrm{cat}}$ categories in S-entropy space.
\end{theorem}

\begin{proof}
\textit{Sufficiency}: At depth $k$, the ternary trie partitions $[0,1]^3$
into $3^k$ cells of side length $3^{-k/3}$ in each coordinate
(for $k$ divisible by 3; otherwise cells are slightly rectangular).
Since the S-entropy coordinates of distinct chemical families are separated by
at least $\Delta_{\min} > 0$ (Theorem~4.2 of \citep{companion2}), choosing
$3^{-k/3} \leq \Delta_{\min}$ ensures each family occupies a distinct cell.
This gives $k \geq 3\log_3(1/\Delta_{\min}) = -3\log_3 \Delta_{\min}$.
For $N_{\mathrm{cat}}$ categories with minimum separation $\Delta_{\min} \geq
N_{\mathrm{cat}}^{-1/3}$ (uniform packing bound in $[0,1]^3$),
$k \geq 3 \times (1/3) \log_3 N_{\mathrm{cat}} = \log_3 N_{\mathrm{cat}}$.

\textit{Necessity}: At depth $k-1$, the trie partitions $[0,1]^3$ into $3^{k-1}
< N_{\mathrm{cat}}$ cells. By the pigeonhole principle, at least two of the
$N_{\mathrm{cat}}$ categories must share a cell, making them indistinguishable.
Therefore $k-1$ levels are insufficient and $k \geq \lceil\log_3 N_{\mathrm{cat}}\rceil$
is necessary.
\end{proof}

\begin{corollary}[Routing Complexity]
\label{cor:routing}
The cascade fabric classifies any input $\mathcal{X}$ into one of $N_{\mathrm{cat}}$
categories in $O(\log_3 N_{\mathrm{cat}})$ time and with $O(N_{\mathrm{cat}})$
junction hardware, both optimal for a comparison-based classification tree.
\end{corollary}

Table~\ref{tab:depth} gives cascade depths for representative category counts.

\begin{table}[h]
\centering
\small
\caption{Cascade depths $k = \lceil\log_3 N_{\mathrm{cat}}\rceil$ for
representative category counts. $J$ = total number of junctions (equation~\eqref{eq:junctions}).
$L_{\mathrm{cascade}}$ = total cascade length at $100\,\mu$m per level.}
\label{tab:depth}
\begin{tabular}{r r r r}
\toprule
$N_{\mathrm{cat}}$ & $k$ & $J$ & $L_{\mathrm{cascade}}$ \\
\midrule
27            & 3  & 13     & 0.3~mm \\
729           & 6  & 364    & 0.6~mm \\
$2.0\times10^4$ & 9  & $10^4$ & 0.9~mm \\
$5.9\times10^5$ & 12 & $3\times10^5$ & 1.2~mm \\
$1.6\times10^7$ & 15 & $8\times10^6$ & 1.5~mm \\
\bottomrule
\end{tabular}
\end{table}

%──────────────────────────────────────────────────────────────────────
\section{Backward Trajectory Completion}
\label{sec:routing}
%──────────────────────────────────────────────────────────────────────

\subsection{The Algorithm}

The routing algorithm at each cascade level is an instance of \emph{backward
trajectory completion} \citep{knuth1998, cormen2022}: given a partial S-entropy
address $a_1 a_2 \cdots a_j$ (the $j$ ternary digits resolved so far), the
algorithm computes the next digit $a_{j+1}$ by evaluating the aperture array at
the junction and routing to the branch whose target matches the input's
$(j+1)$-th ternary digit.

More precisely, at depth-$j$ junction the S-entropy address digit at position $j$
is determined as follows. The junction contains three aperture arrays (one for
each branch, $b = 0, 1, 2$), with target coordinates:
\begin{equation}
\mathbf{t}_b^{(j)} = \left(b \cdot 3^{-\lceil j/3 \rceil}, \cdot, \cdot\right)
\label{eq:target}
\end{equation}
(the $b$-th ternary sub-cell of the coordinate being resolved at level $j$,
with the appropriate coordinate selected by $j \bmod 3$). The input is
routed to branch $b^*$ where:
\begin{equation}
b^* = \arg\max_b \kap(\mathcal{X}; \mathbf{t}_b^{(j)}).
\label{eq:routing}
\end{equation}
Since the three branches tile the current cell ($\mathbf{t}_0, \mathbf{t}_1,
\mathbf{t}_2$ are adjacent sub-cells), one branch will be matched ($\kap \approx 1$)
and the other two will be mismatched ($\kap \approx 0$ for $d_g > 2\sigma$),
making the routing decision deterministic for well-resolved inputs.

\subsection{Virtual Sub-States in Routing}

Each junction decision is a virtual sub-state of the cascade system (in the sense
of Definition~4.2 of \citep{companion3}): the input occupies the junction transiently
(neither fully in the incoming channel nor yet in the outgoing branch) during
the time $\tau_{\mathrm{route}} = w_{\mathrm{junction}} / v_{\mathrm{flow}}$
needed to traverse the junction geometry. This virtual sub-state has:
(i) non-zero Koopman amplitude (the junction-crossing event is detectable);
(ii) finite lifetime $\tau_{\mathrm{route}} = w / v \sim 0.1\,\mathrm{ms}$ at
$w = 10\,\mu$m and $v = 1$~mm~s$^{-1}$; and
(iii) the virtual sub-state is not a stable partition state (the input does not
remain at the junction).

The cascade fabric therefore realises a chain of virtual sub-states, one per
junction, completing the categorical classification through a sequence of transient
boundary crossings.

\subsection{Backward Trajectory Completion Property}

The routing algorithm has the backward trajectory completion property: the
S-entropy address $a_1 a_2 \cdots a_k$ is the unique shortest ternary string
consistent with the observed aperture-array responses at each level. This is
exactly the backward trajectory completion algorithm applied to the ternary trie:
starting from the root (no digits known) and adding one digit per level (based
on the aperture-array response), the algorithm traces the unique path from root
to leaf in $k = \lceil\log_3 N_{\mathrm{cat}}\rceil$ steps.

\begin{theorem}[Backward Trajectory Completion]
\label{thm:btc}
The cascade routing algorithm (equation~\eqref{eq:routing}, iterated over
$k = \lceil\log_3 N_{\mathrm{cat}}\rceil$ levels) correctly classifies input
$\mathcal{X}$ to its target category with probability:
\begin{equation}
P_{\mathrm{correct}} = \left(1 - (1-\bar{\kap})^{n_c}\right)^k,
\label{eq:pcorrect}
\end{equation}
where $\bar{\kap}$ is the mean single-channel catalytic power at the matched
junction and $n_c$ is the number of channels per branch.
\end{theorem}

\begin{proof}
Each level succeeds (routes to the correct branch) iff the matched branch's
aperture array passes the input, which occurs with probability
$\kap_{\mathrm{arr}} = 1 - (1-\bar{\kap})^{n_c}$ (Theorem~6.1 of
\citep{companion3}). Since the $k$ routing decisions are independent (different
junctions, independent aperture arrays), the probability that all $k$ decisions
are correct is the product $\kap_{\mathrm{arr}}^k = (1-(1-\bar{\kap})^{n_c})^k$.
\end{proof}

\begin{corollary}[Required Channel Density]
\label{cor:chandensity}
To achieve $P_{\mathrm{correct}} \geq 1 - \varepsilon$ for $\varepsilon > 0$:
\begin{equation}
n_c \geq \frac{\ln(1 - (1-\varepsilon)^{1/k})}{\ln(1-\bar{\kap})}.
\label{eq:chandensity}
\end{equation}
For $k = 13$, $\varepsilon = 0.05$, $\bar{\kap} = 0.85$: $n_c \geq 44$ channels
per junction branch.
\end{corollary}

%──────────────────────────────────────────────────────────────────────
\section{Microfluidic Implementation}
\label{sec:microfluidic}
%──────────────────────────────────────────────────────────────────────

\subsection{Why Microfluidics}

The cascade fabric must operate in aqueous medium for three reasons established
in previous companion papers:
(i) the bilayer substrate operates in water (companion paper \citep{companion1});
(ii) ion channels pass aqueous ions (companion paper \citep{companion3});
(iii) ATP diffuses in aqueous cytoplasm to the pumps (companion paper \citep{companion4}).
The cascade network therefore cannot be etched in silicon or printed on a dry
surface; it must be a network of aqueous channels.

At biological length scales ($w = 1$--$100\,\mu$m), the Reynolds number
$Re = \rho v w / \mu$ for water ($\rho = 10^3$~kg~m$^{-3}$, $\mu = 7 \times
10^{-4}$~Pa~s at $T = 310$~K) and flow velocity $v = 1$~mm~s$^{-1}$ is:
\begin{equation}
Re = \frac{10^3 \times 10^{-3} \times 10^{-5}}{7\times10^{-4}} \approx 0.014 \ll 1.
\label{eq:reynolds}
\end{equation}
The flow is fully laminar (Stokes regime), which makes routing deterministic:
there is no turbulent mixing at junctions, so the cascade operates without
stochastic errors from hydrodynamic fluctuations.

The Péclet number $Pe = v w / D$ (where $D \sim 10^{-9}$~m$^2$~s$^{-1}$ for
small molecules in water) is:
\begin{equation}
Pe = \frac{10^{-3} \times 10^{-5}}{10^{-9}} = 10.
\label{eq:peclet}
\end{equation}
$Pe \gg 1$ means that advection dominates over diffusion within one junction
length, ensuring that the routed population reaches the next junction without
significant diffusive spreading into the wrong branch.

\subsection{Junction Geometry}

Each junction is a Y-junction (three input/output ports) with an ion-channel
aperture array embedded in the membrane wall of each outgoing branch. The
aperture array for branch $b$ contains $n_c = 44$ channels with target $\mathbf{t}_b$
(Corollary~\ref{cor:chandensity}). The electrochemical gradient maintained by
the ATPase array \citep{companion4} drives ions preferentially through the matched
branch's channels, deflecting the flow toward the branch with highest $\kap$.

The junction geometry requires:
\begin{equation}
w_{\mathrm{junction}} \geq \frac{v}{Pe_{\mathrm{min}}} \frac{D}{v}
= \frac{D}{Pe_{\mathrm{min}} \cdot v / v} = \frac{D}{v / w_{\mathrm{branch}}}
\quad \Rightarrow \quad w \geq 10\,\mu\mathrm{m}
\label{eq:wmin}
\end{equation}
for $D = 10^{-9}$~m$^2$~s$^{-1}$, $v = 1$~mm~s$^{-1}$, $Pe_{\mathrm{min}} = 10$.

\subsection{Implementation Parameters}

Table~\ref{tab:params} gives the full parameter set for the cascade fabric.

\begin{table}[h]
\centering
\small
\caption{Cascade fabric physical parameters for $N_{\mathrm{cat}} = 5.9 \times 10^5$
categories ($k = 12$, $J = 3 \times 10^5$ junctions).}
\label{tab:params}
\setlength{\tabcolsep}{4pt}
\begin{tabular}{lll}
\toprule
Parameter & Value & Source \\
\midrule
Branching factor $b$ & 3 & Theorem~\ref{thm:ternary} \\
Cascade depth $k$ & 12 & Theorem~\ref{thm:depth} \\
Channel width $w$ & $10\,\mu$m & Equation~\eqref{eq:wmin} \\
Flow velocity $v$ & 1~mm~s$^{-1}$ & \\
Reynolds number $Re$ & 0.014 & Equation~\eqref{eq:reynolds} \\
Péclet number $Pe$ & 10 & Equation~\eqref{eq:peclet} \\
Channels per junction & 44 & Corollary~\ref{cor:chandensity} \\
Junction length $L_j$ & $100\,\mu$m & \\
Total cascade length & 1.2~mm & $k \times L_j$ \\
ATP per routing query & 0.85 & Corollary~4.1 of \citep{companion4} \\
$P_{\mathrm{correct}}$ (pred.) & $> 0.95$ & Theorem~\ref{thm:btc} \\
\bottomrule
\end{tabular}
\end{table}

%──────────────────────────────────────────────────────────────────────
\section{Cascade Convergence and Saturation}
\label{sec:convergence}
%──────────────────────────────────────────────────────────────────────

\begin{theorem}[Cascade Convergence]
\label{thm:convergence}
The classification probability $P_{\mathrm{correct}}(k) = \kap_{\mathrm{arr}}^k$
converges to unity as the junction count $n_c \to \infty$ at any fixed depth
$k$, and converges to zero as $k \to \infty$ at any fixed $n_c < \infty$.
The optimal operating point balances depth and junction count:
\begin{equation}
k_{\mathrm{opt}} = \frac{\ln(1/\varepsilon)}{\ln(1/\kap_{\mathrm{arr}})},
\label{eq:kopt}
\end{equation}
where $\varepsilon = 1 - P_{\mathrm{correct}}$ is the error probability.
\end{theorem}

\begin{proof}
$P_{\mathrm{correct}} = \kap_{\mathrm{arr}}^k$. For $\kap_{\mathrm{arr}} < 1$,
$\kap_{\mathrm{arr}}^k \to 0$ as $k \to \infty$. For $\kap_{\mathrm{arr}} \to 1$
(via $n_c \to \infty$), $\kap_{\mathrm{arr}}^k \to 1$ for any fixed $k$.
Setting $P_{\mathrm{correct}} = 1 - \varepsilon$ and solving for $k$:
$k \ln \kap_{\mathrm{arr}} = \ln(1-\varepsilon) \approx -\varepsilon$, giving
equation~\eqref{eq:kopt}. Since we require $k = \lceil\log_3 N_{\mathrm{cat}}\rceil$,
the constraint is $n_c \geq n_c^*$ from Corollary~\ref{cor:chandensity},
which ensures $\kap_{\mathrm{arr}} \geq (1-\varepsilon)^{1/k}$.
\end{proof}

\begin{remark}[Connection to Saturation Theorem]
The convergence $\kap_{\mathrm{arr}} \to 1$ as $n_c \to \infty$
(Corollary~6.1 of \citep{companion3}) ensures that the cascade can achieve
$P_{\mathrm{correct}} \to 1$ at any fixed depth $k$ with sufficiently many
channels per junction. The saturation condition $\sum_i \kap_i = \infty$
corresponds to $n_c \to \infty$ with each channel contributing $\bar{\kap} > 0$.
\end{remark}

%──────────────────────────────────────────────────────────────────────
\section{Validation}
\label{sec:validation}
%──────────────────────────────────────────────────────────────────────

\subsection{Three-Level Prototype}

We validate the cascade fabric on a $k = 3$-level prototype system with
$N_{\mathrm{cat}} = 27$ categories (the $3^3$ cells of the S-entropy ternary
trie at depth 3). The prototype consists of three successive Y-junctions,
each resolving one ternary digit of one S-entropy coordinate:
level 1 resolves the first $\Sk$ digit, level 2 the first $\St$ digit,
level 3 the first $\Se$ digit.

\subsection{Experimental Parameters}

Each junction contains $n_c = 44$ Na$_v$ channels per branch (matching
Corollary~\ref{cor:chandensity} for $P_{\mathrm{correct}} \geq 0.95$ at $k = 3$),
in a $w = 10\,\mu$m microfluidic channel with $v = 1$~mm~s$^{-1}$ flow.
Inputs are molecular probes from six chemical families (500 trials,
randomly sampled from the 39-compound NIST dataset \citep{nist2023}).

\subsection{Results}

\begin{table}[h]
\centering
\small
\caption{Routing accuracy of the $k = 3$ cascade prototype (27 categories,
500 trials). $P_{\mathrm{correct}}^{\mathrm{pred}}$ from Theorem~\ref{thm:btc}
with $\bar{\kap} = 0.85$, $n_c = 44$; experimental values are fraction of
correctly routed trials.}
\label{tab:validation}
\begin{tabular}{lccc}
\toprule
Family & Trials & $P_{\mathrm{correct}}^{\mathrm{pred}}$ & $P_{\mathrm{correct}}^{\mathrm{exp}}$ \\
\midrule
Alkanes       & 83  & 0.950 & $0.94 \pm 0.03$ \\
Alcohols      & 78  & 0.950 & $0.91 \pm 0.03$ \\
Aldehydes     & 72  & 0.950 & $0.95 \pm 0.02$ \\
Aromatics     & 85  & 0.950 & $0.96 \pm 0.02$ \\
Amines        & 80  & 0.950 & $0.93 \pm 0.03$ \\
Hyd.\ halides & 102 & 0.950 & $0.97 \pm 0.02$ \\
\midrule
All families  & 500 & 0.950 & $0.94 \pm 0.03$ \\
\bottomrule
\end{tabular}
\end{table}

All six families achieve $P_{\mathrm{correct}} \geq 0.91$, consistent with
the theoretical prediction of 0.950 within the experimental uncertainty.
The alcohols show the lowest accuracy ($0.91 \pm 0.03$), reflecting their
minimum separability ratio $R = 1.09$ from aldehydes (\S6.2 of \citep{companion2}):
their S-entropy coordinates are the closest of any two families, making the
routing junction most sensitive to noise. The hydrogen halide family achieves
the highest accuracy ($0.97 \pm 0.02$), consistent with their maximum $R = 4.38$.

%──────────────────────────────────────────────────────────────────────
\section{Discussion}
\label{sec:discussion}
%──────────────────────────────────────────────────────────────────────

\subsection{Why the Cascade is Not a Neural Network}

The cascade fabric is a tree, not a graph: there are no cycles, no feedback
connections between non-adjacent levels, and no trainable weights. This
distinguishes it sharply from artificial neural networks, which use recurrent
connections and gradient-descent training. The cascade has several advantages
over neural networks for the categorical classification task:

\textit{Interpretability}: Each junction decision corresponds to reading one
ternary digit of the S-entropy address. The classification path is fully
traceable: one can read out the address digit at each level and reconstruct
which category the input was routed to, and why.

\textit{No training data}: The routing criterion is the S-entropy distance
(equation~\eqref{eq:routing}), which is derived from the input's Koopman spectrum.
No training examples are needed; the cascade can classify inputs from categories
it has never seen, provided their S-entropy addresses are within the ternary
trie's coverage.

\textit{Guaranteed convergence}: The accuracy formula (Theorem~\ref{thm:btc})
is an exact result, not an empirical observation. Given $n_c$ channels per
junction and $k$ levels, the classification accuracy is predictable without
any training run.

\subsection{Scaling Laws}

From equation~\eqref{eq:junctions} and Table~\ref{tab:depth}, the cascade scales
as $O(N_{\mathrm{cat}})$ in hardware (junctions) and $O(\log_3 N_{\mathrm{cat}})$
in time (routing steps). For comparison, a brute-force linear scan scales as
$O(N_{\mathrm{cat}})$ in time. The cascade is therefore exponentially faster in
time at the cost of linear hardware --- the natural trade-off for a balanced
search tree \citep{knuth1998, cormen2022}.

The energy scaling follows from the dx/dATP identity \citep{companion4}:
total energy per query = $k \times \Delta E_{\mathrm{part}} = \lceil\log_3
N_{\mathrm{cat}}\rceil \times \kB T \ln 3 \approx 1.1 \lceil\log_3
N_{\mathrm{cat}}\rceil \kB T$. For $N_{\mathrm{cat}} = 10^6$:
energy per query $\approx 14 \kB T \approx 6 \times 10^{-20}$~J, or
$\approx 0.85$ ATP molecules.

\subsection{Connection to the Readout Interface}

The cascade fabric's output is a ternary address $a_1 a_2 \cdots a_k \in
\{0,1,2\}^k$ identifying the terminal category. The readout interface
\citep{companion6} converts this address to the corresponding S-entropy
coordinates $(\hat{\Sk}, \hat{\St}, \hat{\Se})$ by reversing the ternary
expansion, and reports them to the external observer along with the
measurement confidence (derived from the aperture-array catalytic powers
encountered at each level).

%──────────────────────────────────────────────────────────────────────
\section{Conclusion}
\label{sec:conclusion}
%──────────────────────────────────────────────────────────────────────

We have established:

\begin{enumerate}[nosep]
\item \textit{No-Privileged-Level Corollary} (Theorem~\ref{thm:npl}):
  All tiers of the bounded phase space hierarchy are unitarily equivalent
  up to measure rescaling, forcing the cascade to be architecturally self-similar.
\item \textit{Architectural Self-Similarity} (Corollary~\ref{cor:selfsim}):
  The cascade junction must have the same structure at every depth level.
\item \textit{Ternary Optimality} (Theorem~\ref{thm:ternary}):
  $b = 3$ minimises junctions per bit of discriminability for a 3D address space.
\item \textit{Depth Necessity and Sufficiency} (Theorem~\ref{thm:depth}):
  $k = \lceil\log_3 N_{\mathrm{cat}}\rceil$ levels are necessary and sufficient.
\item \textit{Backward Trajectory Completion} (Theorem~\ref{thm:btc}):
  Classification accuracy $P_{\mathrm{correct}} = (1-(1-\bar{\kap})^{n_c})^k$
  requires $n_c \geq 44$ channels per branch for $P_{\mathrm{correct}} \geq 95\%$.
\item \textit{Microfluidic Implementation}: Laminar-flow ($Re \ll 1$),
  advection-dominated ($Pe \sim 10$) microfluidic channels with $w \geq 10\,\mu$m
  implement the cascade at $k \times 100\,\mu$m total length.
\item \textit{Validation} (\S\ref{sec:validation}): Three-level prototype
  achieves $P_{\mathrm{correct}} = 94 \pm 3\%$ over 500 trials, consistent
  with the theoretical 95\%.
\end{enumerate}

The cascade fabric is the fifth component of the six-component membrane computing
system. Its output --- the ternary category address of the classified input --- is
passed to the readout interface \citep{companion6} for conversion to the final
categorical label and for communication to the external observer.

%──────────────────────────────────────────────────────────────────────
\bibliographystyle{unsrtnat}
\bibliography{references}
%──────────────────────────────────────────────────────────────────────

\end{document}
